---
title: "Gapminder Analysis in R"
author: "Xuan Jiang"
date: 2026-09-21
---

Data: \[Gapminder\] (www.gapminder.org)

```{r}
#| message: false
#| warning: false
# Import Dataframe
library(gapminder)
library(tidyverse)
library(dplyr)
library(ggplot2)

gapminder_subset <- subset(gapminder, country %in% c("Canada", "Japan"))
```

```{r}
# Calculate and select total GDP of Canada and Japan
gapminder_subset <- gapminder_subset |>
  mutate(total_gdp = gdpPercap * pop)

gdp_summary <- gapminder_subset |>
  group_by(country, year) |>
  summarise(total_gdp = sum(total_gdp), .groups = "drop") |>
  mutate(total_gdp_trillion = total_gdp / 10^12) 
```

```{r}
#| label: fig-canada-japan-gdp
#| fig-cap: "Total GDP (in trillions USD) of Canada and Japan over time."
# Plot Total GDP of Canada and Japan over time
ggplot(gdp_summary, aes(x = year, y = total_gdp_trillion, color = country)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(
    x = "Year",
    y = "Total GDP (trillions USD)",
    color = "Country",
    title = "figure 1: \nTotal GDP: Canada vs. Japan"
  ) +
  theme_minimal()
```

```{r}
#| label: fig-canada-japan-gdppercap
#| fig-cap: "GDP per capita of Canada and Japan over time."
# Plot GDP per capita of Canada and Japan over time
ggplot(gapminder_subset, aes(x = year, y = gdpPercap, color = country)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(
    x = "Year",
    y = "GDP per Capita",
    color = "Country",
    title = "figure 2: \nGDP per Capita: Canada vs. Japan"
  ) +
  theme_minimal()
```

I used gapminder dataset to compare the economic growth between Canada and Japan.

First I multiplied the GDP per-capita by population to obtain Canada and Japan's total GDP in trillion USD. Then I visualized the comparison of total GDP and GDP per-capita of both countries over years. From figure 1 , I found Japan's total GDP has been growing rapidly and always greater than Canada's total GDP.

However, this does not imply that the economy in Japan grew faster than Canada once I look into figure 2: the comparison of GDP per capita. Both countries had similar GDP per-capita and Canada's GDP per-capita overtook Japan's for most of the time.
